
---
title: Problem Set week_one
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE, echo=FALSE}
# Load libraries and source extra code
library(tidyverse)
library(readxl)
library(RTutor)


# Options for rendering data frames
# If you knit to a Word docx file, try
# 
# data.frame.theme="word" 
# 
# (needs RStudio > 1.2.1)
# 
# You can also set the options like
# table.max.cols as chunk options
# Makes sense if there are too many, too wide
# columns in some chunks

RTutor::set.knit.print.opts(data.frame.theme="code", table.max.rows=25, table.max.cols=NULL, round.digits=5, signif.digits=8)


# continue knitting even if there is an error
knitr::opts_chunk$set(error = TRUE) 
```


## Exercise Overview

<h1>Week 1 - Descriptive Statistics & Data Visualization using ggplot2</h1>


<h4>Author: Xander Posner, Epi/Biostat MPH '20</h4>
<h4>Check out the <code>source code</code> for this project on <a href="https://github.com/posnerab/RTutorial-Week1", target = _blank>Github</a></h4>



<h3>Tools Used in this Tutorial</h3>

<p>In this tutorial, we will: </p>

<ul>
<li></li>
<li></li>
<li></li>
</ul>


<h3>R Packages Used in this Tutorial</h3>

<ul>
<li><code>"tidyverse"</code></li>
<li><code>"readxl"</code></li>
</ul>

<p>When practicing on your own in RStudio, you will first need to install each of these packages using the <code>install.packages("")</code> function, putting the name(s) of the package(s) inside the quotation marks like this:</p> 

<div id="block-quotes" class="section level2">
<blockquote>
<p><code>install.packages("tidyverse")</code> OR <code>install.packages(c("tidyverse", "readxl"))</code></p>
</blockquote>
</div>

<p>You <strong>won't</strong> need to <i>install</i> packages more than once on the same computer.</p>

<p>After installing, you must load each package library <strong>separately</strong>, using the <code>library()</code> function, putting the name of each package inside the parentheses without quotation marks like this:</p>

<div id="block-quotes" class="section level2">
<blockquote>
<li><p><code>library(tidyverse)</code></p></li>
<li><p><code>library(readxl)</code></p></li>
</blockquote>
</div>

<p>You <strong>will</strong> need to <i>load</i> packages each time you need to use them, using the <code>library()</code> function whenever you initiate a new R session.</p> 

<p><strong>You don't need to install any packages for this tutorial</strong>. Just press <strong>check</strong> to run any chunk with a <code>library()</code> function in it.</p>


<h3>R Functions Used in this Tutorial</h3>

<ul>
<li><code>library()</code> to load each package library</li>
<li><code>read_excel()</code> to read in Excel spreadsheets</li>
<li><code>head()</code> to view the first few lines of your dataset</li>
<li><code>nrow()</code> to view the number of rows in your dataset</li>
<li><code>summary()</code> to produce summary statistics</li>
<li><code>mean()</code> to find the mean of a given variable</li>
<li><code>sd()</code> to find the standard deviation of a given variable</li>


Click the "Go to next exercise..." button to continue.


## Exercise 1 -- Get Started

<p>Some code chunks are finished for you; just press <strong>check</strong> to proceed. Other chunks require you to modify and/or add code before pressing <strong>check</strong> to move through the tutorial.</p>
<p>If you get stuck, use the <strong>hint</strong> in each code chunk first. If you're still having trouble, head to Piazza and post your question.</p>

<strong>Make sure you run each chunk in order!</strong>

<h3>Loading and getting to know your data</h3> 

a) This chunk loads the package libraries, reads in the Excel spreadsheet using the <code>read_excel()</code> function, and assigns the spreadsheet to an object called <code>bmi_adol_men</code> in the global environment.  
```{r "1_a"}
## Don't modify this code; just press check to continue.
library(tidyverse)
library(readxl)
library(epiR)



bmi_adol_men <- read_excel("bmi_young_men.xlsx", col_names = T)
```


b) View just the first few lines of the <code>bmi_adol_men</code> dataset by using the <code>head()</code> function. 
<p><code>head()</code> is a really useful function both when starting analysis and getting acquainted with your dataset or if you just need a quick reminder of the column names.</p>
```{r "1_b_2"}
head(bmi_adol_men)
```

<p>The <code>head()</code> function by default shows you the first six lines, but you can change this if you'd like, using the <code>n = </code> argument inside the <code>head()</code> function.</p> 


c) View the number of rows in your dataset, which here amounts to our sample size, $n$.
```{r "1_c_3"}
nrow(bmi_adol_men)
```

<h3>Summarizing your data</h3> 

<p>R has many ways to view summary statistics. The simplest is the <code>summary()</code> function.</p>

<p>The dollar sign <code>$</code> is used to index or "choose" column names that represent different variables in your dataset with this syntax:</p>

<div id="block-quotes" class="section level2">
<blockquote>
<p><code><"DATASET OBJECT">$<"COLUMN NAME"></code></p>
</blockquote>
</div>

d) Run the chunk below to see what summary statistics the <code>summary()</code> function produces.
```{r "1_d_4"}
summary(bmi_adol_men$bmi)
```



***

### Award: Summary
You summarized it! Click 'Go to next exercise...' to continue.

***




## Exercise 2 -- Calculating Confidence Intervals for a Mean

a) Find the mean and standard deviation of <code>bmi_adol_men$bmi</code> using the <code>mean()</code> and <code>sd()</code> functions.

<p>The first is done for you; mean BMI is already assigned to an object called <code>mean_bmi</code>. Modify the code below it to assign the object <code>std_dev_bmi</code> to the standard deviation of <code>bmi_adol_men$bmi</code> using the <code>sd()</code> function.</p>
```{r "2_a_5"}
bmi_adol_men <- read_excel("bmi_young_men.xlsx", col_names = T)
mean_bmi <- mean(bmi_adol_men$bmi)
mean_bmi
std_dev_bmi <- "<<<<<<<YOUR CODE HERE>>>>>>>"
std_dev_bmi
std_dev_bmi <- sd(bmi_adol_men$bmi)
```


b) Find the standard error of the sample estimator of the population mean using this formula: $$\frac{σ}{\sqrt{n}}$$ where $σ$ is the standard deviation and $n$ is your sample size.
```{r "2_b_6"}
se_bmi <- "<<<<<<<YOUR CODE HERE>>>>>>>"
se_bmi
se_bmi <- sd(bmi_adol_men$bmi)/sqrt(50)
se_bmi
```



Quiz: What would be your alpha for a 95% CI?

- 0.95 [ ]

- 5 [ ]

- 0.025 [ ]

- 0.05 [x]



Quiz: What is the probability, p, or area under the t-distribution curve for a 95% CI?

- 0.95 [ ]

- 0.025 [ ]

- 0.975 [x]

- 0.05 [ ]




c) Modify the code below to find the critical $t^*$ value for a 95% confidence interval (CI) using the <code>qt()</code> function.
```{r "2_c_7"}
t_star_95 <- qt(p = "<<<<<<<YOUR CODE HERE>>>>>>>", 
                df = "<<<<<<<YOUR CODE HERE>>>>>>>", 
                lower.tail = "<<<<<<<TRUE OR FALSE>>>>>>>")
t_star_95
t_star_95 <- qt(p = 0.975, df = 49, lower.tail = TRUE)
t_star_95

```

    


Quiz: What would be your alpha for a 90% CI?

- 0.10 [x]

- 10 [ ]

- 0.05 [ ]

- 0.9 [ ]





Quiz: What would be your p in the <code>qt()</code> function above for a 90% CI?

- 0.9 [ ]

- 0.975 [ ]

- 0.95 [x]

- 0.1 [ ]


